Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-04-07 MAIN.tex
[MG] just started with notation
[MGa] stochastic framework

\newcommand{\cedpage}{\newpage{\thispagestyle{empty}\cleardoublepage}}
\usepackage{algorithm,algpseudocode}
\usepackage{tikz,tikz-3dplot}\usetikzlibrary{arrows.meta,angles}

@@

\author{Martino Gulisano\textsuperscript{a}, Matteo Masoni\textsuperscript{a}, and \name{Marco Gabiccini,\textsuperscript{a,*}
\thanks{* Contact: Marco Gabiccini. Email: marco.gabiccini@unipi.it}
}\affil{
\textsuperscript{a} Dipartimento di Ingegneria Civile e Industriale,	Universit\`{a} di Pisa, Pisa, Italy.}
}

>>>>

\newcommand{\cedpage}{\newpage{\thispagestyle{empty}\cleardoublepage}}
\usepackage{algorithm,algpseudocode}
\usepackage{tikz,tikz-3dplot}
\usetikzlibrary{arrows.meta,angles}

@@

\author{Martino Gulisano\textsuperscript{a}, Matteo Masoni\textsuperscript{a}, and \name{Marco Gabiccini\textsuperscript{a,*}
\thanks{* Contact: Marco Gabiccini. Email: marco.gabiccini@unipi.it}
}\affil{
\textsuperscript{a} Dipartimento di Ingegneria Civile e Industriale,	Universit\`{a} di Pisa, Pisa, Italy.}
}

### 2025-04-07 stochastic_vehicle_dynamics.tex
[MGa] stochastic framework

Due to the symmetry of the probability density function (pdf) with respect to $\bmu(t)$, it is possible to represent the time evolution of the pdf as: i) the deterministic evolution of the mean $\bmu(t)$
\begin{align}
\dbmu(t) = f(\bmu(t), \bu(t)),
\end{align}
and ii) the time evolution of the state covariance matrix $\bP(t)$ along $\bmu(t)$, which can be expressed as
\begin{align}
\dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t),\quad \bP(0) = \bP_0.
\end{align}
The \emph{initial} state covariance $\bP_0$ is assumed known and $\bA(t)$ is the classical shorthand for the Jacobian along the mean trajectory $\bmu(t)$, that is $\bA(t)=\bJ(\bmu(t),\bu(t))$, where $\bJ(\bx, \bu) = \frac{\pd f(\bx, \bu)}{\pd \bx}$.

\begin{align}
\dbx = \begin{bmatrix}
              \dot{\bq}\\
              \dot{\bq}_v
             \end{bmatrix} =
             \begin{bmatrix}
              \bq_v\\
              \text{ABA}\text{dyn}(\bq, \bq_v, f_{xa}, f_{xb}, \delta, \bw),
             \end{bmatrix} = F(\bx).
	\label{eq:ABA_dyn}
\end{align}

>>>>

Due to the symmetry of the probability density function (pdf) with respect to $\bmu(t)$, it is possible to represent the time evolution of the pdf as: i) the deterministic evolution of the mean $\bmu(t)$
\begin{align}\label{eq:meandynamics}
\dbmu(t) = f(\bmu(t), \bu(t)),
\end{align}
and ii) the time evolution of the state covariance matrix $\bP(t)$ along $\bmu(t)$, which can be expressed by the Lyapunov matrix differential equation
\begin{align}\label{eq:dP}
\dbP(t) = \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t),\quad \bP(0) = \bP_0 = \bP_0^T.
\end{align}
In~\eqref{eq:dP}, $\bP_0$ is the \emph{initial} state covariance and $\bA(t)$ is the usual shorthand notation for the Jacobian along the mean trajectory $\bmu(t)$, that is $\bA(t)=\bJ(\bmu(t),\bu(t))$, where $\bJ(\bx, \bu) = \frac{\pd f(\bx, \bu)}{\pd \bx}$.

As can be readily verified by differentiation~\cite{Gajic:LyapunovMatrixEquation:2010}, the analytical solution of~\eqref{eq:dP} has the form
\begin{align}\label{eq:P_STM}
\bP(t) = \bPhi(t,t_0) \bP_0 \bPhi(t,t_0)+\int_{t_0}^{t} \bPhi(t,\tau) \bQ(\tau) \bPhi^T(t,\tau) \dd \tau \quad \bP(0)=\bP_0,
\end{align}
where $\bPhi(t,t_0)$ is the \emph{state transition matrix}. This matrix encodes the evolution from $t_0$ to $t$ of a perturbation w.r.t. to a fiducial trajectory, the mean $\bmu(t)$. In symbols, $\de \bx(t) = \bPhi(t,t_0) \de \bx(t_0)$, with $\de \bx(\bullet) =\bx(\bullet) - \bmu(\bullet) $. In turn, the evolution of $\bPhi(t,t_k)$ from a generic $t_k$ to $t$ is driven by the following differential equation
\begin{align}\label{eq:STM}
\dbPhi(t,t_k) = \bA(t)\bPhi(t,t_k), \quad \bPhi(t_k, t_k) = \bI.
\end{align}
Even if in geometric mechanic and orbital mechanic books, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for a statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}, in view of using collocation integrators in stochastic trajectory planning, we proceed more directly using~\eqref{eq:meandynamics} and~\eqref{eq:dP}.
\begin{align}
\dbx = \begin{bmatrix}
              \dot{\bq}\\
              \dot{\bq}_v
             \end{bmatrix} =
             \begin{bmatrix}
              \bq_v\\
              \text{ABA}\text{dyn}(\bq, \bq_v, f_{xa}, f_{xb}, \delta, \bw),
             \end{bmatrix} = F(\bx).
	\label{eq:ABA_dyn}
\end{align}
